%
% $Id$
%
% $Log$
% Revision 1.3  2005/10/28 14:25:13  sliver
% *** empty log message ***
%
% Revision 1.2  2005/10/26 14:35:52  sbecker
% Started section on provided interfaces
%
%

\section{Interface Relations}

An interface protocol is used to characterise a set of allowed call sequences
sent by a client to a server. However the actual meaning of the protocol of the
interface depends on its relationship to a specific component. TODO: Soll das
wirklich hier hin? Provided und Required Roles/Ports?


\subsection*{Interpretations of Interface Relations}
%---------- Das vielleicht hier rein??????
So far, we identified three different meanings of the specified required
interfaces of a type:
\begin{enumerate}
\item	A required interface can be used, but additional interfaces can be
added.
\item	Only the listed required interfaces can be used.
\item	A required interface has to be used in a predefined way and certain call
sequences have to be executed. 
\end{enumerate}

Required interfaces of the provided-type are not considered explicitly. However,
if they are specified, their meaning corresponds to the first case in the list,
since, for provided-types, the use of external services is not limited. The
second case corresponds to the complete-type. The usage of external services is
limited to the required interfaces specified for the component and no further
interfaces can be used by the component. The last case expresses additional
component requirements. For example, we could specify that all information has
to be stored in a database. So, the interface of the database must be called
using certain call sequences.
%-----------
